---
title: "titanic"
output: html_document
---

Analysis of  the relationships between class, gender, fare, and survival on
The Titanic in the form of a narrative.

Load tidyverse library
```{r}
library(tidyverse)
```

Read in data from train.csv and assign it to variable called "train"

```{r}
train <- read_csv('train.csv')
train
```


We are going to observe three dimensional relationship from our train table and
describe the meaning of each relation we create.

Relationships should be expressed in three ways: as a three-dimensional bar plot
of counts as a three-dimensional bar plot of proportions, and then and as table
of proportions

First of all, for each pair of variables among class, gender, and fare, we are
going to analyze the relationship of survival to that pair count

#################Comparison###########
####Survival with Class & Gender######
######################################

I want to create two separate graphs of male and female, and will observe the
relation between class and Gender with survival.

```{r}
first_plot_count <-
  train %>%
  ggplot(aes(Pclass, fill = factor(Survived))) +
  geom_bar() +
  facet_wrap(~ Sex)
```

# Observation: 
If we compare female and male graph, it is clear that in all three classes,
survival rate for female is much higher than male.

Female graph: We see that almost all the female passenger survive from class1
and class2, and almost half of the female passenger of class3 survived.

Male graph: We see that male passenger in class1 survived the most. Male
passenger in class2 has second highest survival rate and class3 has the least.

Just by looking at male graph, it looks like male survived less in class2 than
class3, but it should be considered that in class3 there are three times more
passengers than class2. We can assure which class has more survival rate if we
look at the graph of proportion of the survival rate.

After my prior observation, I want to see the actual proportion of survival
rate in two graphs for both male and female which will confirm my visual
explanation.

```{r}
first_plot_prop <-
  train %>%
  ggplot(aes(Pclass, fill = factor(Survived))) +
  geom_bar(position = 'fill') +
  facet_wrap(~ Sex)
```

# Observation:
The survival proportion of female passengers in all three classes is higher than
the survival proportion of male passengers. Or we can say in each class more
female passengers have survived than male passengers.

I want to create a table which will give me the exact proportion of survival
rate for male and female.

```{r}
first_table_of_prop <-
  train %>%
  count(Pclass, Survived, Sex) %>%
  spread(Sex, n) %>%
  mutate(female_prop = (female / (female + male)) *100) %>% 
  mutate(male_prop = (male / (female + male)) *100)
```

# Observation:
The proportion of survived female passenger in class1 is 66.9%
The proportion of survived female passenger in class2 is 80.5%
The proportion of survived female passenger in class3 is 60.5%

The proportion of survived male passenger in class1 is 33.1%
The proportion of survived male passenger in class2 is 19.5%
The proportion of survived male passenger in class3 is 39.5%

###########Comparison###########
##Survival with Gender & Class##
################################

I want to create three graphs of all classes, and will observe the relation
between Gender and class with survival.

```{r}
second_plot_count <-
  train %>%
  ggplot(aes(Sex, fill = factor(Survived))) +
  geom_bar() +
  facet_wrap(~ Pclass)
```

# Observation:
People is class1 has higher survival rate than people in class2 and finally
survival rate for class3 is the least. 

Female graph: Most of the female in class1 and class2 have survived where only
half of the female survived in class3. 

Male graph: Less than 50% of male survived in class1. In class2, one fourth of
male survived. In class3 survival rate is the least.


After my prior observation, I want to see the actual proportion of survival rate 
in all three classes which will confirm my visual explaination.
```{r}
second_plot_prop <-
  train %>%
  ggplot(aes(Sex, fill = factor(Survived))) +
  geom_bar(position = 'fill') +
  facet_wrap(~ Pclass)
```

# Observation:
For both graphs we see that the proportion of survival is higher in class1, 
then class2, and finally class3.

In class1, survival proportion of female is 2 times higher than survival
proportion of male. 

In class2, survival proportion of female is almost 4 times higher than
survival proportion of male.

In class3, survival proportion of female is almost 3 times higher than
survival proportion of male.

Among three classes, survival proportion of female is much higher than the
survival proportion of female.

# table prop
```{r}
second_table_of_prop <-
  train %>%
  count(Pclass, Survived, Sex) %>%
  spread(Survived, n) %>%
  mutate(prop_passenger_in_class = (`1` / (`1` + `0`)) * 100)
```

# Observation:

# In class1: 
Proportion of female who survived is 96.8%
Proportion of male who survived is 36.9%

# In class2: 
Proportion of female who survived is 92.1%
Proportion of male who survived is 15.7%

# In class3: 
Proportion of female who survived is 50%
Proportion of male who survived is 13.5%


###########Comparison###########
###Survival with Fare & Class###
################################

I want to create three separate graphs of all three classes, and will observe
the relation between fare and class with survival.
```{r}
third_plot_count <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  ggplot(aes(factor(Fare), fill = factor(Survived))) +
  geom_bar() +
  facet_wrap(~ Pclass)
```

# Observation:
I have broke fare in three intervals, so that it will be easy for us to
visualize the graph.

# In class1: passengers who paid the fare 
between $0 to $8.66: none of them survived.
between $8.66 to $26: a small portion of them survived.
between $26 to $512: looks like almost 75% of the passengers suvived.

# In class2: passengers who paid the fare 
between $0 to $8.66: none of them survived.
between $8.66 to $26: 50% of the passengers suvived.
between $26 to $512: looks like almost 65% of the passengers suvived.

# In class3: passengers who paid the fare 
between $0 to $8.66: looks like one fourth of them survived.
between $8.66 to $26: less than 50% of the passengers suvived.
between $26 to $512: looks like almost 15-20% of the passengers suvived.

After my prior observation, I want to see the actual proportion of survival
rate among three graphs for all three classes which will confirm my visual
explaination.
```{r}
third_plot_prop <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  ggplot(aes(factor(Fare), fill = factor(Survived))) +
  geom_bar(position = 'fill') +
  facet_wrap(~ Pclass)
```

# Observation:
Fare $0 to $8.66: In class1 and class2, no passengers were survived who paid
the fare between $0 to $8.66; while in class3 we see almost 20% passengers
survived by paying the same amount of fare.

Fare $8.66 to $26: In class2 passenger who paid the fare between $8.66 to $26
has the highest survival percentage than class1 and class3.

Fare $26 to $512: In class1 passenger who paid the fare between $26 to $512
has the highest survival percentage than class2 and class3.


I want to create a table which will give me the exact proportion of survival
rate for all passengers in three classes.
```{r}
third_table_of_prop <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  count(Pclass, Fare, Survived) %>%
  spread(Survived, n) %>%
  mutate(`1` =
           case_when(
             is.na(`1`) ~ 0L,
             !is.na(`1`) ~ `1`
           )) %>%
  mutate(survived_prop = (`1` / (`1` + `0`)) * 100)
```

# Observation:
In class1, passenger who paid $0 to $8.66: 0% survival proportion.
In class2, passenger who paid $8.66 to $26: 33.3% survival proportion.
In class3, passenger who paid $0 to $8.66: 65.7% survival proportion.

In class1, passenger who paid $0 to $8.66: 0% survival proportion.
In class2, passenger who paid $8.66 to $26: 46% survival proportion.
In class3, passenger who paid $0 to $8.66: 59% survival proportion.

In class1, passenger who paid $0 to $8.66: 20.6% survival proportion.
In class2, passenger who paid $8.66 to $26: 35% survival proportion.
In class3, passenger who paid $0 to $8.66: 15.4% survival proportion.


###########Comparison###########
###Survival with Class & Fare###
################################

I want to create three separate graphs of three fare intervals, and will
observe the relation between class and fare with survival.
```{r}
fourth_plot_count <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  ggplot(aes(Pclass, fill = factor(Survived))) +
  geom_bar() +
  facet_wrap(~ Fare)
```

# Observation:
passengers who paid $0 to $8.66: In class1 and class2, none of them survived,
however in class3 more than 50 passengers survived paying the same amount of
fare.

Passengers who paid $8.66 to $26: In class1, a very small number of passengers
survived, however in class2 and class3 more than 50 passengers survived paying the same
amount of fare.

Passengers who paid $26 to $512: In class1, two third passengers survived. In
class2, more than 50% passengers survived. And in class3 it has the least
survival proportion even though the passengers paid the same amount of fare.

We can say that passengers who paid more has the better survival chance.

After my prior observation, I want to see the actual proportion of survival
rate among three graphs for all three fare intervals which will confirm my
# visual explaination.
```{r}
fourth_plot_prop <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  ggplot(aes(Pclass, fill = factor(Survived))) +
  geom_bar(position = 'fill') +
  facet_wrap(~ Fare)
```

# Observation:
Fare interval from $0 to $8.66: In class1 and class2, survival proportion is
0%. In class3, survival proportion is near 23%.

Fare interval from $8.66 to $26: In class1, survival proportion is around 30%.
In class2, survival proportion is around 45%. In class3, survival proportion
is around 32%.

Fare interval from $26 to $512: In class1, survival proportion is around 65%.
In class2, survival proportion is around 57%. In class3, survival proportion
is the least and it's around 15%.

I want to create a table which will give me the exact proportion of survival
rate for all passengers with their different fare.
```{r}
fourth_table_of_prop <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  count(Fare, Survived, Pclass) %>%
  spread(Survived, n) %>%
  mutate(`1` =
           case_when(
             is.na(`1`) ~ 0L,
             !is.na(`1`) ~ `1`
           )) %>%
  mutate(surv_prop = (`1` / (`1` + `0`)) * 100)
```

# Observation:

# Fare interval $0 to $8.66:
In class1, survival proportion is 0%
In class2, survival proportion is 0%
In class3, survival proportion is 20.6%

# Fare interval $8.66 to $26:
In class1, survival proportion is 33.3%
In class2, survival proportion is 46%
In class3, survival proportion is 35%

# Fare interval $26 to $512:
In class1, survival proportion is 65.7%
In class2, survival proportion is 59%
In class3, survival proportion is 15.4%

###########Comparison##########
##Survival with Fare & Gender##
###############################

I want to create two separate graphs of male and female, and will
observe the relation between fare and gender with survival.
```{r}
fifth_plot_count <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  ggplot(aes(factor(Fare), fill = factor(Survived))) +
  geom_bar() +
  facet_wrap(~ Sex)
```

# Obsevation:
In the female graph, we see that more than 100 female passengers survive who
paid between $26 to $512. Around 80 female passengers survived who paid
between $8.66 to $26. More than 30 female passengers survived who paid between
$0 to $8.66.

In the male graph, we see that around 50 male passengers survive who
paid between $26 to $512. Male passengers who paid the fare from $0 to $8.66
and $8.66 to $512 has the same count for suvival.

Paying the same amount of fare, female passengers survived more than the male
passengers.


After my prior observation, I want to see the actual proportion of survival
rate between the two graphs for three fare intervals by which I can confirm my
visual explaination.
```{r}
fifth_plot_prop <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  ggplot(aes(factor(Fare), fill = factor(Survived))) +
  geom_bar(position = 'fill') +
  facet_wrap(~ Sex)
```

# Observation:
We see that the survival rate is growing as the fare is increasing; however
there is a huge difference in between male and female graph base on survival.


I want to create a table which will give me the exact proportion of survival
rate for male and female passengers in these three fare intervals.
```{r}
fifth_table_of_prop <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  count(Sex, Fare, Survived) %>%
  spread(Survived, n) %>%
  mutate(prop = `1` / (`1` + `0`) * 100)
```

# Observation:
In female graph: Survival rate is growing as the fare is increasing. Also
survival rate is higher than death rate.

In male graph: Survival rate is growing as the fare is increasing, but
survival rate is less lower than death rate.

###########Comparison##########
##Survival with Gender & Fare##
###############################

I want to create graphs for three fare interval, and will
observe the relation between fare and gender with survival.
```{r}
sixth_plot_count <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  ggplot(aes(Sex, fill = factor(Survived))) +
  geom_bar() +
  facet_wrap(~ Fare)
```

# Obsevation:
Fare interval $0 to $8.66: Out of 50+ female passengers, more than half of
them survived, while out of 250+ male passengers the survival count is less
than female one.

Fare interval $8.66 to $26: This is the reflection of first case; however the
male passenger survival count is higher than the previous one.

Fare interval $26 to $512: Out of almost 135 female passengers, more than 100
of them survived, while out of almost 160 male passengers the survival count
is a little over 50.

After my prior observation, I want to see the actual proportion of survival
rate among three fare intervals with gender by which I can confirm my visual
explaination.
```{r}
sixth_plot_prop <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  ggplot(aes(Sex, fill = factor(Survived))) +
  geom_bar(position = 'fill') +
  facet_wrap(~ Fare)
```

# Observation:
We see that the female's survival rate in all three fare interval is higher
than the male survival rate.

By paying fare between $0 to $8.66, more than 50% of female passengers
survived while male survived less than 10%.
By paying fare between $8.66 to $26, close to 70% of female passengers
survived while male survived little over 15%.
By paying fare between $26 to $512, more than 75% of female passengers
survived while male survived less than 30%.

I want to create a table which will give me the exact proportion of survival
rate for male and female passengers in these three fare intervals.
```{r}
sixth_table_of_prop <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  count(Fare, Survived, Sex) %>%
  spread(Survived, n) %>%
  mutate(`1` =
           case_when(
             is.na(`1`) ~ 0L,
             !is.na(`1`) ~ `1`
           )) %>%
  mutate(prop = (`1` / (`1` + `0`)) * 100)
```

# Observation:
Fare paid between $0 to $8.66: Female survived 63%; male survived 10.6%
Fare paid between $8.66 to $26: Female survived 69%; male survived 17%
Fare paid between $26 to $512: Female survived 84%; male survived 33.5%


##########Comparison##########
###Gender with Fare & Class###
##############################

I want to create graphs for all three classes and break the fare in three fare
interval, and will observe the relation between fare and class with gender.
```{r}
seventh_plot_count <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  ggplot(aes(factor(Fare), fill = factor(Sex))) +
  geom_bar() +
  facet_wrap(~ Pclass)
```

# Observation:

In class1: passengers who paid $0 to $8.66, all of them are male.
passengers who paid $8.66 to $26, most of them are male.
passengers who paid $26 to $512, more than half are male.

In class2: passengers who paid $0 to $8.66, all of them are male.
passengers who paid $8.66 to $26, more than half are male.
passengers who paid $26 to $512, male and female proportion is 50/50.

In class3: passengers who paid $0 to $8.66, almost out of 300 passengers
nearly 235 are male.
passengers who paid $8.66 to $26, and $26 to $512; more than half are male.

```{r}
seventh_plot_prop <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  ggplot(aes(factor(Fare), fill = factor(Sex))) +
  geom_bar(position = 'fill') +
  facet_wrap(~ Pclass)
```

# Observation:

# In class1: 
Passenger who paid $0 to $8.66: all of them are male.
Passenger who paid $8.66 to $26: almost 65% of them are male. 
Passenger who paid $26 to $512: there are almost same amount of male and female
passenger.


```{r}
seventh_table_of_prop <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  count(Pclass, Fare, Sex) %>%
  spread(Sex, n) %>%
  mutate(`female` =
           case_when(
             is.na(`female`) ~ 0L,
             !is.na(`female`) ~ `female`
           )) %>%
  mutate(prop_female = (`female` / (`female` + `male`) * 100)) %>% 
  mutate(prop_male = (`male` / (`female` + `male`) * 100))
```

# Observation:

# In class1: 
Interval [$0 to $8.66]: Proportion of male is 100%, proportion of female is 0%.
Interval ($8.66 to $26]: Proportion of male is 66.7%, proportion of female is 33.3%
Interval ($26 to $512]: Proportion of male is 54.9%, proportion of female is 45.1%

# In class2: 
Interval [$0 to $8.66]: Proportion of male is 100%, proportion of female is 0%.
Interval ($8.66 to $26]: Proportion of male is 59%, proportion of female is 41%
Interval ($26 to $512]: Proportion of male is 51.3%, proportion of female is 48.7%

# In class3: 
Interval [$0 to $8.66]: Proportion of male is 81.8%, proportion of female is 18.2%.
Interval ($8.66 to $26]: Proportion of male is 51%, proportion of female is 49%
Interval ($26 to $512]: Proportion of male is 61.5%, proportion of female is 38.5%

###########Comparison#########
###Fare with Gender & Class###
##############################

I want to create graphs for all three classes and break the fare in three fare
interval, and will observe the relation between gender and class with fare.
```{r}
eighth_plot_count <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  ggplot(aes(factor(Sex), fill = factor(Fare))) +
  geom_bar() +
  facet_wrap(~ Pclass)
```

# Observation

# In class1:
Almost all the female and male passenger paid the highest fare to get in to
class1. Also there are a small portion of male passengers who paid $0 to $8.66
and $8.66 to $26 and still got in to class1.

# In class2:
There are nearly 20 male and female passengers whose fare is in the highest
interval but got in to class2. There are large portion of male and female
passenger whose fare interval is $8.66 to $26. Also a small portion of male
passenger got in to class2 by paying the least fare interval.

# In class3:
We see that there is a decent number of passengers who paid the fare from
highest interval but got in class3. There is same amount of male and female
passengers in class3 who paid the fare from $8.66 to $26. Almost 4 times more
male than female in class3 whose fare interval is $0 to $8.66.


```{r}
eighth_plot_prop <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  ggplot(aes(factor(Sex), fill = factor(Fare))) +
  geom_bar(position = 'fill') +
  facet_wrap(~ Pclass)
```

# Observation:
The higher the fare that people paid, the higher the class they are in.
However, there are passengers who paid the least amount for fare, and they are
in higher class.


```{r}
eighth_table_of_prop <-
  train %>%
  mutate(Fare = cut_number(Fare, 3)) %>%
  count(Sex, Fare, Pclass) %>%
  spread(Fare, n) %>% 
  mutate(`[0,8.66]` = 
           case_when(
             is.na(`[0,8.66]`) ~ 0L,
             !is.na(`[0,8.66]`) ~ `[0,8.66]`
           )) %>%
  mutate(prop_fare1 = (`[0,8.66]` / (`[0,8.66]` + `(8.66,26]` + `(26,512]`)) * 100) %>%
  mutate(prop_fare2 = (`(8.66,26]` / (`[0,8.66]` + `(8.66,26]` + `(26,512]`)) * 100) %>%
  mutate(prop_fare3 = (`(26,512]` / (`[0,8.66]` + `(8.66,26]` + `(26,512]`)) * 100)
```

# Observation:
# In class1: 
female passenger who paid the fare between $0 to $8.66 is 0% and male 4.92%
female passenger who paid the fare between $8.66 to $26 is 2.13% and male 3.28%
female passenger who paid the fare between $26 to $512 is 97.9% and male 91.8%

# In class2: 
female passenger who paid the fare between $0 to $8.66 is 0% and male 5.56%
female passenger who paid the fare between $8.66 to $26 is 75% and male 75.9%
female passenger who paid the fare between $26 to $512 is 25% and male 18.5%

# In class3: 
female passenger who paid the fare between $0 to $8.66 is 37.5% and male 69.7%
female passenger who paid the fare between $8.66 to $26 is 48.6% and male 21%
female passenger who paid the fare between $26 to $512 is 13.9% and male 9.22%

# End of code